Sperm should evolve to make female meiosis fair.

### 2014-07-28 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb finalized changes for resub 7/27

{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
O\"{o}genesis is inherently susceptible to such conflicts
because alleles compete for inclusion in the  product of female meiosis that will be transmitted to the egg. 
Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
In this light, many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
That females  of many animal species do not complete meiosis until after
     fertilization appears to run counter to this interpretation
     because this delay provides an opportunity for sperm-acting alleles to
     meddle with the outcome of female meiosis and help like-alleles drive in heterozygous females. 
Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
	in fact, sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. 
These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals to date.  
Rather than providing an opportunity for sperm collaboration in female meiotic drive,
        the `fertilization requirement'  indirectly protects females from
        meiotic drivers by providing sperm an opportunity to
        suppress drive. 
%	and suggest that the requirement of fertilization for the
%%        completion of female meiosis
%        potentially represents a mechanism employed by females 
%	 to ensure a fair meiosis.
\newpage

\section*{Introduction}
Despite the apparent unity of the organism, `selfish' alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), may increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and the organism.
Meiotic drivers observed in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
conflict -- the selfish benefits and the associated
pleiotropic fitness costs of drive sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}. \newline

@@

Before doing so, it is worth considering whether there is any evidence that sperm could (or do), influence female meiosis. 
It is becoming
        increasingly clear that sperm bring a wide variety of RNA and proteins into the
        egg \citep{Miller:05}. 
 Some of these have known functions, for example, in most
        animal species sperm -- not eggs -- are
        responsible for the transmission of the centriole, a vital
        component of the mitotic machinery for the zygote \citep{Schatten:94}. 
Detailed functional studies and analyses of paternal effect mutations in model
systems further highlight that sperm-transmitted products have a wide-range of functions in
 egg activation, completion of syngamy, zygotic development, and the resumption
 and successful completion of
 female meiosis \cite[e.g.][]{Yasuda:95,Loppin:05,Miller:01, McNally:05,Churchill:03}. 
 For example, in \emph{C. elegans}, premature deployment of the sperm aster 
	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
However, the function of many of the products the sperm brings into the
        egg is completely unknown and 
      these products vary widely over species \citep{Karrbook:09}. 
It seems quite plausible that sperm-based products, and
	hence sperm haplotype or paternal genotype could influence 
	various aspects of female meiosis that occur after fertilization. 
\newline

@@

Numerous lines of evidence suggest that female meiotic drive is a common and important 
	evolutionary force, and therefore the opportunity for sperm to influence female drive likely had applications to many animals.   
While research to date has identified a few extreme cases of female meiotic 
	drive in the small number model systems systematically
        studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
	rapid evolution of the basic components of the meiotic machinery (e.g. centromeres, telomeres, etc \dots) 
	suggest consistent selection on female meiotic drivers and suppressors of meiotic drive in many animal species 
	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}. 
 We expect that over the next decade the spread of sequencing to a
         range of systems will reveal many more female meiotic drive
         systems, however, carefully characterizing them will still
         remain a challenging task. 
\newline

@@

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	\yb{and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.}  
Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg likely evolved as a mechanism 
	to enforce an equitable transfer of genetic material.

@@

%%%sperm-acting facilitator
When the modifier of drive arises on the drive background (i.e. in coupling phase),
	is tightly linked to the driver, and enhances drive 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
        Model 4 in the Appendix and File S1 ). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, 
	the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2  ). \newline

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, creating a
        third allele $B^{-}$ (Figure
          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       it does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background ( i.e. in repulsion phase, 
	creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Equations \ref{A-} and
        \ref{unlinked}).  %%DELETED FIG REF HERE. 
 We note that the evolution of sperm-acting drive  suppressors unlinked to a driver (Model 6)
 	is both qualitatively and quantitatively similar to the evolution  of a female-acting drive 
	suppressor (Model 6$^\prime$ -- compare Equations \ref{unlinked} and \ref{femunlinked}). 
        \newline

The sperm acting drive suppressing allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive. 
\newline

@@

The theory presented herein provides a strong prediction -- 
	for organisms in which female  meiosis is not completed until after fertilization, 
	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
Contrary to expectations of a naive verbal ``green-beard'' model, 
	our results are consistent with the observation that female meiosis 
	is more fair when fertilized by sperm bearing the drive allele in 
	two of the best described cases of female meiotic drive in 
	animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).  
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
We further predict that segregation in F1 females backcrossed to parental species will likely be biased with a deficit 
	of transmission of the paternal species allele from the F1  female.   
We note that while these predictions follow straightforwardly from the
	theory presented above, tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline

\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although 
        opportunities to test this hypothesis may be limited because 
        %this effect may be weaks as such
        lineages may only persist for a short time). 
Our results also suggests that 
%Another   hypothesis that our work suggests is we might expect 
        that phylogenetic variation in the stage of female meiosis when
        fertilization occurs may influence the prevalence of
        female meiotic drive.
For example, in taxa where females complete meiosis
	before fertilization, centromeric drive may
	be more common than in species which sperm interact with eggs arrested  in MI, 
	because in the prior case sperm-based modifiers can only intercede during the $2^\text{nd}$, but not $1^\text{st}$ meiotic division. }
As a potential test of this hypothesis the speed of centromere
	turnover could be compared in species in which sperm iterate with eggs paused at MI and MII 
	(assuming the pace of centromere turnover serves as a proxy for the efficacy of MI drive). 
\newline

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}.
However, given the existence of female meiotic arrest, 
	\gc{and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} 
	the nature of the meiotic arrest and interactions between
        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
	and selfish drivers evolving in concert to overcome these adaptations. }
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }

@@

\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
\caption{Invasion analysis for a self-promoting female meiotic drive allele with
  recessive costs (selection coefficient $s$), showing the region of
  bistability. The colors, and the thin dashed contours, indicate the
  frequency the allele must be above to invade the population (note
  that these alleles reach fixation conditional on invading). In the
  white area the allele cannot invade ($x \geq1$), in the solid red
  area the allele can invade and fix when rare. In the left panel we
  show the results obtained by a grid search using the recursion, on
  the right we show the approximation obtained assuming that HWE
  holds. }  
\label{Bistab_homozyg_cost_fig}
\end{figure}

>>>>

{\bf Abstract:}
Genomic conflicts arise when an allele gains an evolutionary advantage at a cost to organismal fitness. 
O\"{o}genesis is inherently susceptible to such conflicts
because alleles compete for inclusion in the  product of female meiosis that will be transmitted to the egg. 
Alleles that distort meiosis in their favor (i.e. meiotic drivers) often decrease organismal fitness,
	and therefore indirectly favor the evolution of mechanisms to suppress meiotic drive. 
In this light, many facets of  o\"{o}genesis and gametogenesis have been 
	interpreted as mechanisms of protection against genomic outlaws. 
That females  of many animal species do not complete meiosis until after fertilization, 
     appears to run counter to this interpretation,
     because this delay provides an opportunity for sperm-acting alleles to
     meddle with the outcome of female meiosis and help like alleles drive in heterozygous females. 
Contrary to this perceived danger, the population genetic theory presented herein suggests that, 
	in fact, sperm nearly always evolve to increase the fairness of female
        meiosis in the face of genomic conflicts. 
These results are consistent with the apparent sperm-dependence of the best characterized female meiotic drivers in animals.   
Rather than providing an opportunity for sperm collaboration in female meiotic drive,
        the `fertilization requirement'  indirectly protects females from
        meiotic drivers by providing sperm an opportunity to
        suppress drive. 
%	and suggest that the requirement of fertilization for the
%%        completion of female meiosis
%        potentially represents a mechanism employed by females 
%	 to ensure a fair meiosis.
\newpage

\section*{Introduction}
Despite the apparent unity of the organism, `selfish' alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), can increase in frequency even if it entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and the organism.
Meiotic drivers observed in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
conflict -- the selfish benefits and the associated
pleiotropic fitness costs of drive sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, the structure of centromere machinery \citep{Malik2002a,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}. \newline

@@

Before doing so, we highlight the evidence that sperm can (or do), influence female meiosis. 
It is becoming increasingly clear that sperm bring a wide variety of RNA and proteins into the egg \citep{Miller:05}. 
 Some of these have known functions, for example, in most
        animal species, sperm -- not eggs -- are
        responsible for the transmission of the centriole, a vital
        component of the mitotic machinery for the zygote \citep{Schatten:94}. 
Detailed functional studies and analyses of paternal effect mutations in model
systems further highlight that sperm-transmitted products have a wide-range of functions in
 egg activation, completion of syngamy, zygotic development, and the resumption
 and successful completion of
 female meiosis \cite[e.g.][]{Yasuda:95,Loppin:05,Miller:01, McNally:05,Churchill:03}. 
 For example, in \emph{C. elegans}, premature deployment of the sperm aster 
	disrupts MII meiotic segregation in the egg, leading to a triploid zygote \citep{McNally2012}.
However, the function of many of the products the sperm brings into the
        egg is completely unknown and 
      these products vary widely over species \citep{Karrbook:09}. 
It seems quite plausible that sperm-based products, and
	hence sperm haplotype or paternal genotype could influence 
	various aspects of female meiosis that occur after fertilization. 
\newline

@@

Numerous lines of evidence suggest that female meiotic drive is a common and important 
	evolutionary force, and therefore the opportunity for sperm to influence female drive is likely relevant to many animals.    
While research to date has identified a few extreme cases of female meiotic 
	drive in the small number model systems systematically
        studied \citep{Agulnik1990, Fishman2008, Hiatt2003,Novitski1951, Pardo-ManuelDeVillena2001c},
	rapid evolution of the basic components of the meiotic machinery (e.g. centromeres, telomeres, etc \dots) 
	suggest consistent selection on female meiotic drivers and suppressors of meiotic drive in many animal species 
	 \citep[e.g. ][]{Anderson2008, Anderson2009, Axelsson2010, Malik2009a}. 
 We expect that over the next decade the spread of sequencing to a
         range of systems will reveal many more female meiotic drive
         systems, however, carefully characterizing them will still
         remain a challenging task. 
\newline

@@

Because female meiotic drive is likely a common force with predictably negative effects on organismal fitness,  
	and because sperm have ample opportunity to influence female drive,  
	we develop population genetic models to address the the expected influence of sperm on female drive.  
We first focus on models in which `self-promoting' alleles in
  sperm facilitate drive of like alleles during gametogenesis in heterozygous females. 
These models show that such sperm-acting alleles 
	have more difficulty invading a population than do traditional meiotic drivers, 
	 and under most circumstances, cannot be maintained as a balanced polymorphism.
Because self-promoting drivers are unlikely to create a sustained genomic conflict, 
	female meiosis will have little incentive to evolve resistance to them.
We then examine models in which a novel sperm-acting allele modifies the efficacy of a polymorphic meiotic driver. 
Such models universally reveal that sperm-acting drive modifiers are favored only if they suppress drive. 
These results highlight the fact that the interests of sperm and
maternal  genomes' are often aligned, 
as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
Thus, there is little selective benefit to females in preventing sperm to influence female meioses, 
	and in fact, females eschewing this delay would potentially lose sperm assistance in the suppression of meiotic drivers.  
Given the wide-spread requirement of fertilization for the completion
	of  female meiosis, various features of the interaction between sperm 
	and egg may function to enforce an equitable transfer of genetic material.

@@

%%%sperm-acting facilitator
When the modifier of drive arises on the drive background (i.e. in coupling phase),
	is tightly linked to the driver, and enhances drive 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$
        heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}A2, and
        Model 4 in the Appendix and File S1). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, 
	the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and File S2). \newline

@@

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, creating a
        third allele $B^{-}$ (Figure
          \ref{Eggsperm_3_allele_cartoon}A3, Model 4 in Appendix). This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       it does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes. This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}B1, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (i.e. in repulsion phase, 
	creating a third allele $A^{-}$, Figure \ref{Eggsperm_3_allele_cartoon}A4, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Equations \ref{A-} and
        \ref{unlinked}).  %%DELETED FIG REF HERE. 
 We note that the evolution of sperm-acting drive  suppressors unlinked to a driver (Model 6)
 	is both qualitatively and quantitatively similar to the evolution  of a female-acting drive 
	suppressor (Model 6$^\prime$ -- compare Equations \ref{unlinked} and \ref{femunlinked}). 
        \newline

The sperm acting drive suppressing allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon}B2). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive, and can do so as effectively (or more effectively) than female-acting drive modifiers. 
\newline

@@

The theory presented herein provides a strong prediction -- 
	for organisms in which female  meiosis is not completed until after fertilization, 
	sperm will act to suppress female drive at the stage at which they can influence meiosis. 
%Therefore, for example, when fertilized by drying sperm, we expect meiosis to be fairer when  females are fertilized by 
Contrary to expectations of a naive verbal ``green-beard'' model, 
	our results are consistent with the observation that female meiosis 
	is fairer when fertilized by sperm bearing the drive allele in 
	two of the best described cases of female meiotic drive in 
	animals (the \emph{Om} and \emph{In} loci in mice \citep{Agulnik1993,Wu2005}).  
Given the benefit to sperm of hampering female drive, 
	drive-suppressing sperm are often likely to be fixed within a species, 
	making the hypothesis of sperm-acting drive-suppression difficult to test from intra-population crosses. 
However, crosses between populations or species are likely to provide critical tests of our theory -- 
	specifically we predict that female meiosis will be less fair when a species (or population) is 
	fertilized by heterospecific sperm because such sperm have not evolved to counter novel females meiotic drivers. 
We further predict that segregation in F1 females backcrossed to parental species will likely be biased, 
	with a deficit of transmission of the paternal species allele from the F1  female.   
Tthese predictions follow straightforwardly from the
	theory presented above; however, we caution that tests of meiotic drive, 
	and especially sperm-dependent meiotic drive require a high standard
	of evidence  to exclude plausible alternative hypotheses such as genotypic 
	inviability including epistatic maternal X zygotic lethality \citep[e.g. ][]{Sawamura:1993aa}.  \newline

Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although 
        opportunities to test this hypothesis may be limited because 
        %this effect may be weaks as such
        lineages may only persist for a short time). 
Our results also suggests that phylogenetic variation in the stage of female meiosis when
        fertilization occurs may influence the prevalence of  female meiotic drive.
For example, centromeric drive may be more common  in taxa where females complete MI before fertilization, 
	as compared to species in which sperm interact with eggs arrested  in MI, 
	because in the prior case sperm-based modifiers can only intercede during the second, but not the first  meiotic division. 
As a potential test of this hypothesis, the speed of centromere
	turnover could be compared in species in which sperm iteract with eggs paused at MI and MII 
	(assuming the pace of centromere turnover serves as a proxy for the frequency of MI drivers). 
\newline

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}, in fact, we presuppose its existence.
However, given the existence of female meiotic arrest, 
	and that its timing and mechanistic details are variable across species \citep{Masui_book,Karrbook:09} ,
	the nature of the meiotic arrest and interactions between
        sperm and egg may  be molded by selection to reduce the opportunity for female meiotic drive, 
	and counteracted by selfish drivers evolving to overcome these adaptations. 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }

@@

\includegraphics[width = \textwidth]{../../Scripts/homozyg_bistability.eps}
\caption{Invasion analysis for a self-promoting female meiotic drive allele with
  recessive costs (selection coefficient $s$), showing the region of
  bistability. The colors, and the thin dashed contours, indicate the
  frequency the allele must reach, $f^*$ in order to invade the population (note
  that these alleles reach fixation conditional on invading). In the
  white area, the allele cannot invade, in the solid red
  area the allele can invade and fix when rare. In the left panel we
  show the results obtained by a grid search using the recursion, on
  the right we show the approximation obtained assuming that HWE
  holds. }  
\label{Bistab_homozyg_cost_fig}
\end{figure}
